---
title: "p8105_hw1_rh2916"
author: "Rui Huang"
date: "September 19, 2018"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```


## Problem 1

## Create data frame

```{r p8105_hw1_problem1}
problem1_df = tibble(
  vec_numeric = runif(10, min = 0, max = 5),
  vec_logical = (vec_numeric > 2),
  vec_char = c("This", "is", "P8105", "homework1", "problem1", "and", "my", "uni", "is",                   "rh2916"),
  vec_factor = factor(c("Yes", "Yes", "No", "Yes", "No", "Yes", "Yes", "No", "No", "No"))
)
```

## Mean of each variable 

```{r}
mean(problem1_df$vec_numeric)
mean(problem1_df$vec_logical)
mean(problem1_df$vec_char)
mean(problem1_df$vec_factor)
```

## What works and what doesn’t? Why?
The means of numeric variable and logical variable can be derived, and means of character and factor variables cannot be taken. Numeric variables can have mean. Logical variable is stored as 0 (FALSE) or 1(TRUE), which can have mean. Character variables are a string of one or more characters and factor variables contain a set of numeric codes with character-valued levels. Thus, both of them return NA.

## Convert variables from one type to another

```{r eval = FALSE}
as.numeric(problem1_df$vec_logical)
as.numeric(problem1_df$vec_char)
as.numeric(problem1_df$vec_factor)
```

##What happens?
Logical and factor variables can be converted to numeric variables by as.numeric function. For logical variables, "FALSE" is converted to 0 while "TRUE" is converted to 1. For factor variables, elements are converted to 1 and the further integers. Character variables cannot be converted to numeric ones.

```{r}
as.numeric(as.factor(problem1_df$vec_char))
as.numeric(as.character(problem1_df$vec_factor))
```

##What happens?
Character variables can be converted to factor variables, then they can be converted to numeric ones. Factor variables can be converted to character variables, then they can no longer be converted to numeric ones.


## Problem 2

## Create data frame

```{r p8105_hw1_problem2}
set.seed(1000)

problem2_df = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  vector_logical = x + y > 0,
  vector_numeric = as.numeric(vector_logical),
  vector_factor = as.factor(vector_logical)
)
```
## Description of vectors
The size of the dataset is `r dim(problem2_df)[1]`
The mean of x is `r mean(problem2_df$x)` and median of c is `r median(problem2_df$x)`
The proportion of cases for which the logical vector is TRUE is `r sum(problem2_df$vector_logical)/dim(problem2_df)[1]`











